---
title: "Project_1_week-4_V2"
author: "Robert Green"
date: "2024-09-27"
output: html_document
---


The required task is to read a text file with chess tournament results and 
generate a .CSV file with information on all the players.

```{r}

library(stringr)
library(stringi)

```


The file with the chess tournament results is read from GITHUB.
```{r message=FALSE, warning=FALSE}

ratings_tab_from_git <- readLines("https://raw.githubusercontent.com/RGreen-sps/scaling-umbrella/refs/heads/main/Project_1_table.txt")

```


The user will enter a 'Pair Num', a value to select the player of interest.
```{r}


player_pair_Num <- 1
#player_pair_Num <- readline("** Player's Pair Number (enter at Console) ** : ")


```


The entered number is group together with other characters to form a search pattern.
```{r}

pair_num_space_pipe <-  c(" ", player_pair_Num, "\\|")

pair_num_search <- str_flatten(pair_num_space_pipe, collapse = " ")

```


The table is searched with the given pattern and the first row position
containing the player's information is located.
```{r}

row_pos <- grep(pair_num_search, ratings_tab_from_git)

```



The first and second row containing all the necessary information is selected.
```{r}

selected_player_row <- ratings_tab_from_git[row_pos]

selected_player_row_2nd_line <- ratings_tab_from_git[row_pos + 1]

```


From the first row, the player's name and 'Total Pts' are extracted
```{r}

selected_player_name       <- substr(selected_player_row, 9, 40)

selected_player_name       <- str_remove_all(selected_player_name, " {2,}")

selected_player_total_pts  <- substr(selected_player_row,  42, 44)

selected_player_total_pts  <- as.double(selected_player_total_pts)

```


From the second row, the players State and rating is extracted. The rating is
converted to an integer.
```{r}
selected_player_state      <- substr(selected_player_row_2nd_line,  4,  5)

selected_player_rating_pre <- substr(selected_player_row_2nd_line, 22, 26)

selected_player_rating_pre <- as.integer(selected_player_rating_pre)
```


The extracted information along with a new field (average) is assigned 
to a data frame. Other players information will be appended to this data frame.
```{r}
average <- 0L

player_s_info_df <- ""


player_s_info_df <- data.frame(selected_player_name, selected_player_state, selected_player_total_pts,
                          selected_player_rating_pre, average)

player_s_info_df
```


All game results, round-1 to round-7 will be saved to a vector. Each element of
the vector will contain the player's number that identifies the player and the 
outcome of the game. For these calculations, only, a win, draw, or loss outcome 
will be used in further calculations.
```{r}

round_1_to_7_vec <- c(substr(selected_player_row,  48, 52),
                      substr(selected_player_row,  54, 58),
                      substr(selected_player_row,  60, 64),
                      substr(selected_player_row,  66, 70),
                      substr(selected_player_row,  72, 76),
                      substr(selected_player_row,  78, 82),
                      substr(selected_player_row,  84, 88))

```



The for loop will examine each element for a win(W), draw(D), or loss(L), 
at which point the selected player's first row and second row will be selected
from the ratings table('ratings_tab_from_git') and, as before, the necessary 
data are extracted and appended to the 'player_s_info_df' data frame.
```{r for loop}

for(i in 1:length(round_1_to_7_vec)) { #vector containing game outcome
  win_draw_lose <- substr(round_1_to_7_vec[i], 1, 1)
  
  #identification of game results
  if(win_draw_lose == "W" | win_draw_lose == "D" | win_draw_lose == "L") {
    player_pair_Num <- str_extract(round_1_to_7_vec[i], "[(0-9)]{1,}")
    
    #selected players number paired with search pattern
    pair_num_space_pipe <-  c(" ", player_pair_Num, "\\|")
    pair_num_search <- str_flatten(pair_num_space_pipe, collapse = " ")
    
    #row position located and selected
    row_pos <- grep(pair_num_search, ratings_tab_from_git) 
    selected_player_row <- ratings_tab_from_git[row_pos]
    selected_player_row_2nd_line <- ratings_tab_from_git[row_pos + 1]
    
    #player's information extracted
    selected_player_name <- substr(selected_player_row, 9, 40)
    selected_player_name <- str_remove_all(selected_player_name, " {2,}")
    
    selected_player_total_pts  <- substr(selected_player_row,  42, 44)
    selected_player_total_pts  <- as.double(selected_player_total_pts)
    
    selected_player_state      <- substr(selected_player_row_2nd_line,  4,  5)
    
    selected_player_rating_pre <- substr(selected_player_row_2nd_line, 22, 26)
    selected_player_rating_pre <- as.integer(selected_player_rating_pre)
    
    #player's information combined and added to list
    player_info_list <- list(selected_player_name,
                          selected_player_state,
                          selected_player_total_pts,
                          selected_player_rating_pre,
                          average)
    
    #end of iteration, player's information added to the data frame
    player_s_info_df <- rbind2(player_s_info_df, player_info_list)
   
  }
}

```


The for loop iterates through each field of the avareage colum of the dataframe and inserts the calculated avearge value.
```{r}

# loop iterates through the last column of the data frame
pre_rate_column <- c(player_s_info_df[, 4])
for(i in 1:length(pre_rate_column)){
  column_total <- sum(player_s_info_df[, 4]) - player_s_info_df[i, 4]
  
  num_of_col_elemts_less_one <- length(pre_rate_column) - 1
  
  average <- column_total / num_of_col_elemts_less_one
  
  player_s_info_df[i, 5] <- as.integer(average)

}

```


The columns names of the data frame is updated as required.
```{r}
colnames(player_s_info_df) <- c("Player’s Name",
                                "Player’s State",
                                "Total Number of Points",
                                "Player’s Pre-Rating",
                                "Average Pre Chess Rating of Opponents")
```


Finally, the results of the opreation is output as a .CSV file to GITHUB.
```{r}

write.csv(player_s_info_df, "chess_stats.csv")

```

